---
title: "DiabetesStudy_pt5"
author: "KAShortt"
date: "July 23, 2017"
output: html_document
---

The final project will analyze a diabetes dataset available courtesy of Dr John Schorling, Department of Medicine, University of Virginia School of Medicine.
	dataset information: http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/diabetes.html
	variable information: http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/Cdiabetes.html
	data: http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/diabetes.xls

## R Markdown
```{r version-and-current-date}
R.version.string
Sys.Date()
```

## Cleaning house

```{r clean-house}
ls()
rm(list=ls()) # This list=ls() option removes everything
```

## use epitools
```{r libraries}
library("epitools")
library("readxl")
```

## Set working directory
```{r set-working}
setwd("C:/Users/kathe/Documents/R class/pt 5 final project")
```

## 1. Compute some summary statistics. Are there any missing values?

## read dataset
```{r read data}
dm <- read_excel("diabetes.xls")
head(dm)
tail(dm)
```

## summary statistics
```{r summary}
overall_summary <- summary(dm)
str(overall_summary)
write.csv(overall_summary, file="summary.csv")
```

## confirm NAs
```{r confirm NAs}
na_count <-sapply(dm, function(y) sum(length(which(is.na(y)))))
data.frame(na_count)

```

## 2. Create a factor for at least one categorical variable.
```{r create binary factor for glycolosolated hemoglobin}

## Glycosolated hemoglobin > 7 consistent with diabetes mellitus
dm$glyhb.num <- as.numeric(dm$glyhb>7)
dm$glyhb.factor <- factor(dm$glyhb.num, levels=c(0,1), labels=c("normal","high"))

## check coding                         
table(dm$glyhb, dm$glyhb.num)
table(dm$glyhb.num,dm$glyhb.factor)
```

## 3. Draw at least one graph appropriate for your data.
```{r barplot}

bmp(filename="barplot1.bmp")
round(100*prop.table(table(dm$glyhb.factor)))
barplot(100*prop.table(table(dm$glyhb.factor)),xlab="Glycosolated hemoglobin", ylab="%",axes=FALSE)
axis(side=2,at=10*(0:10),labels=paste(10*(0:10),"%",sep=""))
title(" ")
dev.off()
```

## 4. Calculate a correlation or an odds ratio (or both) depending on what type of data you have.
```{r correlation}
bio.measures<-c("chol", "hdl", "stab.glu", "glyhb")
phys.measures<-c("age", "height", "weight", "waist", "hip")
round(100*cor(dm[,bio.measures],dm[,phys.measures],use="pairwise.complete.obs"))
```

```{r correlation}
bio.measures<-c("chol", "hdl", "stab.glu")
round(100*cor(dm$glyhb,dm[,bio.measures],use="pairwise.complete.obs"))
```

## 5. Produce at least two other statistics or graphs that you think might be interesting and informative.
```{r remove NAs}
# select variables glyhb, stab.glu
myvars <- c("glyhb","stab.glu")
dm2 <- dm[myvars]

dm3 <- na.omit(dm2)

summary(dm3)
```

```{r scatter plot}
bmp(filename="scatterplot1.bmp")
smooth.curve <- lowess(dm3$glyhb, dm3$stab.glu)
plot(dm$glyhb,dm$stab.glu,
     xlab="Glycosolated Hemoglobin",
     ylab="Stable Glucose",
     pch="+")
title(" ")
lines(smooth.curve)
dev.off()
```

```{r }
## stab.glu by gender
grp.means <- by(dm$stab.glu,dm$gender,mean,na.rm=TRUE)
grp.stdev <- by(dm$stab.glu,dm$gender,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")

## stab.glu by location
grp.means2 <- by(dm$stab.glu,dm$location,mean,na.rm=TRUE)
grp.stdev2 <- by(dm$stab.glu,dm$location,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means2),colon,round(grp.means2,1),
      plus.minus,round(grp.stdev2,1),sep="")

## glyhb by gender
grp.means3 <- by(dm$glyhb,dm$gender,mean,na.rm=TRUE)
grp.stdev3 <- by(dm$glyhb,dm$gender,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means3),colon,round(grp.means3,1),
      plus.minus,round(grp.stdev3,1),sep="")

## glyhb by location
grp.means4 <- by(dm$glyhb,dm$location,mean,na.rm=TRUE)
grp.stdev4 <- by(dm$glyhb,dm$location,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means4),colon,round(grp.means4,1),
      plus.minus,round(grp.stdev4,1),sep="")
```

## 6. Place all the key results into a Word document, PowerPoint presentation, as a PDF file or as an HTML file. Turn it in by email. Provide a brief interpretation in the context of your data set for everything that you produce above.

## Save everything
```{r}
save.image("C:/Users/kathe/Documents/R class/pt 5 final project/DM.RData")
```


End Of File

