---
title: "GardasilStudy_pt2"
author: "KAShortt"
date: "July 4, 2017"
output: html_document
---

The Gardasil vaccine requires three shots in order to be effective. A study conducted at Johns Hopkins looked at how often patients failed to get all three shots. They wanted to see if insurance status, age, and other factors could predict who was at greatest risk for failing to get all three shots.

The data set is available as an Excel spreadsheet at http://www.amstat.org/publications/jse/v19n1/gardasil.xls

and a description of the file is available at http://www.amstat.org/publications/jse/v19n1/gardasil.txt

## R Markdown
```{r version-and-current-date}
R.version.string
Sys.Date()
```

## Cleaning house

```{r clean-house}
ls()
rm(list=ls()) # This list=ls() option removes everything
```

## use epitools
```{r libraries}
library("epitools")
```

## Set working directory

```{r set-working}
setwd("C:/Users/kathe/Documents/R class/pt 2 homework")
```

## read in dataset
```{r read in data}
fn <- "gardasil.csv"
ga <- read.csv(file=fn,header=TRUE)
head(ga)
tail(ga)
```

## 1. Create factors for AgeGroup, Race, Completed, Location.
```{r create factors}
ga$AgeGroup.factor <- factor(ga$AgeGroup, levels=c(0,1), labels=c("child", "adult"))
table(ga$AgeGroup, ga$AgeGroup.factor)

ga$Race.factor <- factor(ga$Race, levels=c(0,1,2,3), labels=c("white", "black", "hispanic", "other/unknown"))
table(ga$Race, ga$Race.factor)

ga$Completed.factor <- factor(ga$Completed, levels=c(0,1), labels=c("no", "yes"))
table(ga$Completed, ga$Completed.factor)

ga$Location.factor <- factor(ga$Location, levels=c(1,2,3,4), labels=c("Odenton", "White Marsh", "JHOC", "Bayview"))
table(ga$Location, ga$Location.factor)
```

## 2. Report which variables have missing data and how many of these values are missing.
```{r missing vlaues}
overall_summary <- summary(ga)
str(overall_summary)
write.csv(overall_summary, file="summary.csv")
```
```{r confirm no NAs}
na_count <-sapply(ga, function(y) sum(length(which(is.na(y)))))
data.frame(na_count)

```
## 3. What proportion of patients completed all three shots?
```{r proportion table}
tb <- table(ga$Completed.factor)
pct <- round(100*prop.table(tb))
n  <- sum(tb)
pct.sign <- "%"
colon <- ":"
slash <- "/"
comma <- ", "
paste(names(tb),colon,tb,slash,n,comma,pct,pct.sign,sep="")
```
## 4. Draw a bar chart showing the percentage of patients at each of the four locations.
```{r barplot}

bmp(filename="barplot1.bmp")
round(100*prop.table(table(ga$Location.factor)))
barplot(100*prop.table(table(ga$Location.factor)),xlab="Location", ylab="%",axes=FALSE)
axis(side=2,at=10*(0:5),labels=paste(10*(0:5),"%",sep=""))
title(" ")
dev.off()
```

## 5. Use a crosstabulation to compare Age to AgeGroup, and Location to LocationType. Are the recodings into AgeGroup and LocationType done properly?
```{r crosstabs}
table(ga$Age,ga$AgeGroup)
table(ga$Location,ga$LocationType)
```

## 6. Create a new variable that combines the race categories into white, and non-white. Calculate the proportion of white patients at each of the four locations.
```{r new race variable}
ga$NonWhite <- as.numeric(ga$Race==1 | ga$Race==2 | ga$Race==3)
table(ga$Race.factor,ga$NonWhite)

ga$NonWhite.factor <- factor(ga$NonWhite, labels=c("white", "nonwhite"))
table(ga$NonWhite,ga$NonWhite.factor)


tb1 <- table(ga$NonWhite.factor,ga$Location.factor)
round(100*prop.table(tb1, margin=1))
```

## 7. Does the likelihood of completing all three shots vary by AgeGroup, Race, or Location? Calculate the appropriate percentages and relative risks.

```{r percentages}
tb2 <- table(ga$AgeGroup.factor, ga$Completed.factor)
round(100*prop.table(tb2, margin=1))

tb3 <- table(ga$Race.factor, ga$Completed.factor)
round(100*prop.table(tb3, margin=1))


tb4 <- table(ga$Location.factor, ga$Completed.factor)
round(100*prop.table(tb4, margin=1))
```

```{r risk ratios}
sink('AgeGroup_vs_Completed_rr.txt')
riskratio(ga$AgeGroup.factor, ga$Completed.factor, rev="columns")
sink()

sink('Race_vs_Completed_rr.txt')
riskratio(ga$Race.factor, ga$Completed.factor, rev="columns")
sink()

sink('Location_vs_Completed_rr.txt')
riskratio(ga$Location.factor, ga$Completed.factor, rev="columns")
sink()
```


## 8. Place all the key results into a Word document or PowerPoint presentation, or as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 2 Homework"

## Save everything
```{r}
save.image("C:/Users/kathe/Documents/R class/pt 2 homework/GardasilHomework.RData")
```

End Of File