---
title: "HousingStudy_pt3"
author: "KAShortt"
date: "July 9, 2017"
output: html_document
---


Look at a data file on housing prices in Albuquerque at
--> http://lib.stat.cmu.edu/DASL/Datafiles/homedat.html

You should cut-and-paste the actual data to notepad and save it with the name housing.txt. You need to read in the file using the read.table function.
For technical reasons, you want the as.is=TRUE argument. Otherwise AGE and TAX are converted to factors which causes all sorts of grief.
There is a mix of categorical and continuous variables. Recall that you use a scatterplot to examine the relationship between two continuous variables, a crosstabls and an odds ratio to examine the relationship between two categorical variables, and a boxplot to examine the relationship between a categorical variable and a continuous variable.

## R Markdown
```{r version-and-current-date}
R.version.string
Sys.Date()
```

## Cleaning house

```{r clean-house}
ls()
rm(list=ls()) # This list=ls() option removes everything
```

## use epitools
```{r libraries}
library("epitools")
```

## Set working directory
```{r set-working}
setwd("C:/Users/kathe/Documents/R class/pt 3 homework")
```

## read in dataset
```{r read in data}
fn <- "housing.txt"
hp <- read.table(file=fn,header=TRUE,as.is=TRUE)
head(hp)
tail(hp)
summary(hp)
```

## 1. Convert the asterisks in the TAX variable to missing. How many missing values are there for AGE and for TAX?


```{r * to NA}
'is.na<-'(hp, hp=="*")
```

```{r convert tax and age to numeric}

hp$TAX.num<-as.numeric(hp$TAX)
hp$AGE.num<-as.numeric(hp$AGE)

```

``` {r new summary}
overall_summary <- summary(hp)
str(overall_summary)
write.csv(overall_summary, file="summary.csv")
```

```{r confirm no NAs}
na_count <-sapply(hp, function(y) sum(length(which(is.na(y)))))
data.frame(na_count)

```

## 2. Create factors for NE, CUST, and COR. Draw bar charts for each of these factors.
```{r create factors}
hp$NE.factor <- factor(hp$NE, levels=c(0,1), labels=c("not", "northeast"))
table(hp$NE, hp$NE.factor)

hp$CUST.factor <- factor(hp$CUST, levels=c(0,1), labels=c("no", "yes"))
table(hp$CUST, hp$CUST.factor)

hp$COR.factor <- factor(hp$COR, levels=c(0,1), labels=c("not", "corner"))
table(hp$COR, hp$COR.factor)
```

```{r barplots}

bmp(filename="barplot1.bmp")
round(100*prop.table(table(hp$NE.factor)))
barplot(100*prop.table(table(hp$NE.factor)),xlab="Location", ylab="%",axes=FALSE)
axis(side=2,at=10*(0:5),labels=paste(10*(0:5),"%",sep=""))
title(" ")
dev.off()

bmp(filename="barplot2.bmp")
round(100*prop.table(table(hp$CUST.factor)))
barplot(100*prop.table(table(hp$CUST.factor)),xlab="Custom Built", ylab="%",axes=FALSE)
axis(side=2,at=10*(0:5),labels=paste(10*(0:5),"%",sep=""))
title(" ")
dev.off()

bmp(filename="barplot3.bmp")
round(100*prop.table(table(hp$COR.factor)))
barplot(100*prop.table(table(hp$COR.factor)),xlab="Corner", ylab="%",axes=FALSE)
axis(side=2,at=10*(0:5),labels=paste(10*(0:5),"%",sep=""))
title(" ")
dev.off()
```

## 3. Find the largest house (biggest SQFT) in the data set. Is the largest house also the most expensive house?
```{r which max house}
hp[which(hp$SQFT==3750),0]
hp[which(hp$PRICE==2150),0]
```

## 4. Calculate frequency counts for FEATS. Are there any houses with the no features? Are there any houses with every possible feature?
```{r freq counts}
table(hp$FEATS)
```

## 5. Evaluate the relationship between PRICE and SQFT using a scatterplot. Include a smooth curve. Do larger houses tend to cost more?
```{r scatter plot}
bmp(filename="scatterplot1.bmp")
smooth.curve <- lowess(hp$SQFT,hp$PRICE)
plot(hp$SQFT,hp$PRICE,
     xlab="Square Feet",
     ylab="Price",
     pch="+")
title(" ")
lines(smooth.curve)
dev.off()
```

## 6. Evaluate the relationship between CUST and PRICE using a boxplot. Calculate the mean price and standard deviation by CUST.
```{r}
bmp(filename="boxplot1.bmp")
boxplot(PRICE~CUST.factor,data=hp,xlab="Customized",ylab="Price")
title(" ")
dev.off()

grp.means <- by(hp$PRICE,hp$CUST.factor,mean,na.rm=TRUE)
grp.stdev <- by(hp$PRICE,hp$CUST.factor,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")
```

## 7. What is the difference in average prices between northeast houses and other houses? What is the difference in average taxes?
```{r }
grp.means <- by(hp$PRICE,hp$NE.factor,mean,na.rm=TRUE)
grp.stdev <- by(hp$PRICE,hp$NE.factor,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")

grp.means <- by(hp$TAX.num,hp$NE.factor,mean,na.rm=TRUE)
grp.stdev <- by(hp$TAX.num,hp$NE.factor,sd,na.rm=TRUE)
colon <- ": "
plus.minus <- "+/-"
paste(names(grp.means),colon,round(grp.means,1),
      plus.minus,round(grp.stdev,1),sep="")
```
If you didnt use as.is=true whe you loaded the data, convert TAX from a factor to a numeric variable: 

## 8. Are custombuilt houses more likely to appear on corner lots? Calculate the percentages and compute a relative risk.

```{r }
tb1<-table(hp$COR.factor, hp$CUST.factor)
round(100*prop.table(tb1, margin=1))

sink('CORvCUST_rr.txt')
riskratio(hp$COR.factor, hp$CUST.factor, rev="columns")
sink()
```

## 9. Place all the key results into a Word document, PowerPoint presentation, as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 3 Homework"

## Save everything
```{r}
save.image("C:/Users/kathe/Documents/R class/pt 3 homework/HOusingHomework.RData")
```

End Of File

