---
title: "SleepInMammals_pt1"
author: "KAShortt"
date: "June 18, 2017"
output: html_document
---

This purpose of this program is to complete the analysis of a "Sleep in Mammals" data set for Intro to R assignment 1.  Data description is available at http://www.statsci.org/data/general/sleep.html and data is available at http://www.statsci.org/data/general/sleep.txt.



## R Markdown
```{r version-and-current-date}
R.version.string
Sys.Date()
```

## Cleaning house

```{r clean-house}
ls()
rm(list=ls()) # This list=ls() option removes everything
```

## Set working directory

```{r set-working}
setwd("C:/Users/kathe/Documents/R class/pt1 homework")
```

## 1. Notice that there is a huge range in body weight. Display the information for the smallest and the largest mammals.

To answer this, read in the data and look at the min and max values in bodywt.


```{r read-summary}
fn <- "C:/Users/kathe/Documents/R class/pt1 homework/sleep.txt"
fd <- read.table(file=fn,header=TRUE,row.names=1)
head(fd)
tail(fd)
overall_summary <- summary(fd)
str(overall_summary)
write.csv(overall_summary, file="summary.csv")
```
Then determin which species correcpond with the min and max values

```{r min-max-BodyWt}
fd[which(fd$BodyWt==0.005),0]
fd[which(fd$BodyWt==6654.000),0]
```

## 2. Which variables have missing data?

here, just look at the NAs in the summary.

## 3. Calculate the mean and standard deviation for TotalSleep.

```{r TotalSleep}
mean(fd$TotalSleep,na.rm=TRUE)
sd(fd$TotalSleep,na.rm=TRUE)
```

## 4. Draw a histogram for the BodyWt variable. Note that this variable is highly skewed. Re-draw the histogram on the log scale. Hint: log(sl$BodyWt,base=10). For exta credit, relabel the axes with values at 0.001, 0.01, etc.

```{r histogram}
bmp(filename="histogram1.bmp")
hist(fd$BodyWt)
dev.off()
```
Plot the log10 transformed data

```{r histogram-modified}

d1 <- log(fd$BodyWt,base=10)
bmp(filename="histogram2.bmp")
hist(d1,main=" ",breaks=c(-3,-2,-1,0,1,2,3,4),xlab="Body Weight in kg",axes=FALSE)
lb <- c("0.001","0.01","0.1","1","10","100","1,000","10,000")
axis(side=1,at=c(-3,-2,-1,0,1,2,3,4),labels=lb)
axis(side=2,at=5*(0:5))
dev.off()
```

## 5. Calculate the ratio of BrainWt to BodyWt and express it as a percentage. Be sure to convert grams to kilograms (or vice versa) before computing the ratio. What animal has the smallest/largest ratio?

```{r BrainWt-vs-BodyWt}
d2<-c(fd$Species,100*(fd$BrainWt/(1000*fd$BodyWt)))
summary(d2)
which.min(d2)
which.max(d2)

```


determine teh species which ahve the min and max ratios

```{r min-max-ratio}
fd[1,]
fd[27,]
```

## 6. Do bigger animals sleep longer or live longer? Show your results using a correlation coefficient.

```{r sleep-vs-live}
SleepLife.measures<-c("TotalSleep","NonDreaming","Dreaming","LifeSpan")
round(100*cor(fd[,"BodyWt"],fd[,SleepLife.measures],use="pairwise.complete.obs"))
```

## 7. Do animals who have high degrees of Predation, Exposure, or Danger sleep less? Show your results using a plot and a trendline.

Correlation between valriables
```{r stress-vs-sleep}
stress.measures<-c("Predation","Exposure","Danger")
sleep.measures<-c("TotalSleep","NonDreaming","Dreaming")
round(100*cor(fd[,stress.measures],fd[,sleep.measures],use="pairwise.complete.obs"))
```

Plot the data and add a trend line

```{r scatter-plots}
bmp(filename="scatterplot1.bmp")
trend.line <- lm(TotalSleep~Predation,data=fd)
plot(fd$Predation,fd$TotalSleep,
     xlab="Predation Index",
     ylab="Total Sleep in Hrs/Day",
     pch="+")
title(" ")
abline(trend.line)
dev.off()

bmp(filename="scatterplot2.bmp")
trend.line <- lm(TotalSleep~Exposure,data=fd)
plot(fd$Exposure,fd$TotalSleep,
     xlab="Exposure",
     ylab="Total Sleep in Hrs/Day",
     pch="+")
title(" ")
abline(trend.line)
dev.off()

bmp(filename="scatterplot3.bmp")
trend.line <- lm(TotalSleep~Danger,data=fd)
plot(fd$Danger,fd$TotalSleep,
     xlab="Danger",
     ylab="Total Sleep in Hrs/Day",
     pch="+")
title(" ")
abline(trend.line)
dev.off()
```

## 8. Place all the key results into a Word document or PowerPoint presentation, or as a PDF file or as an HTML file. Turn it in by email. Use the subject line: "Introduction to R, (your name), Part 1 Homework"

## Save everything
```{r}
save.image("C:/Users/kathe/Documents/R class/pt1 homework/SleepHomework.RData")
```

End Of File